%!TEX root = ../thesis.tex
\begin{abstract}
\section*{Abstract}
In recent years, driven by their unique capabilities, the research on hexapod robots has gained significant traction.
These systems, able to navigate challenging terrain and overcome complex obstacles, hold the promise to advancements in various domains.
While hexpods possess significant potential, real-world applications remain rare, and substantial research is still required to further enhance the current systems. Given the intricacies of hexapod robots as complex mechanical machines, the utilization of simulation software offers numerous advantages.
With this thesis we aim to add to the array of existing simulators by providing a comprehensive, modular and physically accurate model of a hexapod robot using the MATLAB Simulink environment.
The model we develop aims to lay the foundation for future research to build on.
During the course of our work we continuously test the model in numerous simulations, ensuring the functionality of each module and the final product as a whole.
For this purpose, we conceive a simple locomotion controller, enabling omnidirectional locomotion of the hexapod in several different, static gait patterns.
Furthermore, we develop the capabilities to apply Reinforcement Learning to problem of autonomously learning efficient leg coordination strategies, highlighting the versatility of the MATLAB-platform and providing impulses for future research to expand upon these efforts. 

\end{abstract}
